A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-04-24 tidyindex.qmd
Small changes and fixes

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [@wickham_tidy_2014; @wickham_split-apply-combine_2011; @tidymodels; @wang2020; @zhang2024cubble] that "turn ideas into software quickly and faithfully", as envisioned by @chambers1998programming. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->

@@

To be able to test different decision choices for an index, systematically and statistically, the index needs to be broken down into its fundamental building blocks to analyze the contribution and effect of each component. We call this process the *index pipeline*, which are the steps of the data analysis pipeline for the index construction. Such decomposition of index components provides the means to standardize index construction via a pipeline and offers benefits for comparing among indexes, calculating index uncertainty, and assessing index robustness.

@@

Since the success of tidyverse, more packages have been developed to analyze data using the tidy framework for domains specific applications, a noticeable example of which is `tidymodels` for building machine learning models [@tidymodels]. To create a tidy workflow tailored to a specific domain, developers first need to identify the fundamental building blocks to create a workflow. These components are then implemented as modules, which can be combined to form the pipeline. For example, in supervised machine learning models, steps such as data splitting, model training, and model evaluation are commonly used in most workflow. In the `tidymodels`, these steps are correspondingly implemented as package `rsample`, `parsnip`, and `yardstick`, agnostic to the specific model chosen. The uniform interface in tidymodels frees analysts from recalling model-specific syntax for performing the same operation across different models, increasing the efficiency to work with different models simultaneously.

@@

-   makes the computation more transparent, and thus more easily debugged, facilitating reproducibility.
-   allows for rapidly processing new data to check how different features, like outliers, might affect the index value.
-   provides the capacity to measure uncertainty by computing confidence intervals from multiple samples as generated by bootstrapping to original data.
-   enables systematic comparison of surrogate indexes designed to measure the same phenomenon.
-   it may even be possible to automate diagrammatic explanations and documentation of the index.

@@

Unlike variable transformation, scaling maintains the distributional shape of the variable. It includes techniques such as centering, z-score standardization, and min-max standardization and can be expressed as $[x_{j}(s;t) - \alpha]/\gamma$ where $\alpha$ and $\gamma$ are constants . In the Human Development Index (HDI), the three dimensions (health, education, and economy) are converted into the same scale (0-1) using min-max standardization.

@@

The accompanied package, tidyindex, is not intended to offer an exhaustive implementation for all indexes across every domains. Instead, it provides a realization of the pipeline framework proposed in the paper. When adopting the pipeline approach to construct indexes, analysts may consider developing software that can be readily deployed in the cloud for production purposes.

@@

The state of Queensland in Australia frequently experiences natural disaster events such as flood and drought, which can significantly impact its agricultural industry. This example uses daily data from Global Historical Climatology Network Daily (GHCND), aggregated into monthly precipitation, to compute two drought indexes -- SPI and SPEI -- at various time scales and fitted distributions, for  `r length(unique(queensland$id))` stations in the state of Queensland in Australia, spanning from January 1990 to April 2022. This example showcases the basic calculation of indexes with different parameter specifications within the pipeline framework. The dataset used in this example is available in the `tidyindex` package as `queensland` and blow prints the first few rows of the data:

@@

@fig-idx-tour illustrates doing sensitivity analysis for GGGI, for a subset of 16 countries. It presents 6 frames selected from the animation where the weight on the politics dimension is gradually increased, where other dimensions (economy, education, health) decrease correspondingly. Frame 12 present the original index where all the four dimensions receive equal weight. The index values are sorted from highest to lowest, with the Nordic countries (Iceland, Norway, and Finland) and New Zealand leading the rankings. The index values are between 0 and 1, and indicate proportional difference between men and women, with a value of 0.8 indicating women are 80% of the way to equality of these measures. There is a gap in values between these countries and the middle group (Brazil, Panama, Poland, Bangladesh, Kazakhstan, Armenia, and Slovakia), and another big drop to the next group (Pakistan, Iran, Algeria, and Chad). Afghanistan lags much further behind.

@@

To make a simple illustration of sensitivity analysis, we slightly vary the weight for politics, between 0.07 and 0.52, while maintain equal weights among other dimensions. This can be viewed as an animation to examine change in relative index values as a response to the changing weights. This visualization technique, which presents a sequence of data projections, is referred to as a "tour" and the specific kind of tour used here to move between nearby projections is known as a "radial tour" (see @buja2005computational, @wickham_tourr_2011 and @RJ-2020-027 for more details).

>>>>

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring a unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [@wickham_tidy_2014; @wickham_split-apply-combine_2011; @tidymodels; @wang2020; @zhang2024cubble] that "turn ideas into software quickly and faithfully", as envisioned by @chambers1998programming. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->

@@

To be able to test different decision choices for an index, systematically and statistically, the index needs to be broken down into its fundamental building blocks to analyze the contribution and effect of each component. We call this process the *index pipeline*, which are the steps of the data analysis pipeline for index construction. Such a decomposition of index components provides the means to standardize index construction via a pipeline and offers benefits for comparing among indexes, calculating index uncertainty, and assessing index robustness. In addition it provides an unambiguous documentation for the index definition, facilitating reproducibility of results.

@@

Since the success of tidyverse, more packages have been developed to analyze data using the tidy framework for domains specific applications, a noticeable example of which is `tidymodels` for building machine learning models [@tidymodels]. To create a tidy workflow tailored to a specific domain, developers first need to identify the fundamental building blocks to create a workflow. These components are then implemented as modules, which can be combined to form the pipeline. For example, in supervised machine learning models, steps such as data splitting, model training, and model evaluation are commonly used in most workflow. In the `tidymodels`, these steps are correspondingly implemented as packages `rsample`, `parsnip`, and `yardstick`, agnostic to the specific model chosen. The uniform interface in tidymodels frees analysts from recalling model-specific syntax for performing the same operation across different models, increasing the efficiency to work with different models simultaneously.

@@

-   makes the computation more transparent, and thus more easily debugged, facilitating reproducibility.
-   allows for rapidly processing new data to check how different features, like outliers, might affect the index value.
-   provides the capacity to measure uncertainty by computing confidence intervals from multiple samples as generated by bootstrapping the original data.
-   enables systematic comparison of surrogate indexes designed to measure the same phenomenon.
-   it may even be possible to automate diagrammatic explanations and documentation of the index.

@@

Unlike variable transformation, scaling maintains the distributional shape of the variable. It includes techniques such as centering, z-score standardization, and min-max standardization and can be expressed as $[x_{j}(s;t) - \alpha]/\gamma$ where $\alpha$ and $\gamma$ are constants. In the Human Development Index (HDI), the three dimensions (health, education, and economy) are converted into the same scale (0-1) using min-max standardization.

@@

The tidyindex package is not intended to offer an exhaustive implementation for all indexes across all domains. Instead, it provides a realization of the pipeline framework proposed in the paper. When adopting the pipeline approach to construct indexes, analysts may consider developing software that can be readily deployed in the cloud for production purposes.

@@

The state of Queensland in Australia frequently experiences natural disaster events such as flood and drought, which can significantly impact its agricultural industry. This example uses daily data from Global Historical Climatology Network Daily (GHCND), aggregated into monthly precipitation, to compute two drought indexes -- SPI and SPEI -- at various time scales and fitted distributions, for  `r length(unique(queensland$id))` stations in the state of Queensland in Australia, spanning from January 1990 to April 2022. This example showcases the basic calculation of indexes with different parameter specifications within the pipeline framework. The dataset used in this example is available in the `tidyindex` package as `queensland` and below we show the first few rows of the data:

@@

@fig-idx-tour illustrates doing sensitivity analysis for GGGI, for a subset of 16 countries. It presents 6 frames selected from an animation where the weight on the politics dimension is gradually increased, while other dimensions (economy, education, health) decrease correspondingly. Frame 12 present the original index where all the four dimensions receive equal weight. The index values are sorted from highest to lowest, with the Nordic countries (Iceland, Norway, and Finland) and New Zealand leading the rankings. The index values are between 0 and 1, and indicate proportional difference between men and women, with a value of 0.8 indicating women are 80% of the way to equality of these measures. There is a gap in values between these countries and the middle group (Brazil, Panama, Poland, Bangladesh, Kazakhstan, Armenia, and Slovakia), and another big drop to the next group (Pakistan, Iran, Algeria, and Chad). Afghanistan lags much further behind.

@@

To make a simple illustration of sensitivity analysis, we slightly vary the weight for politics, between 0.07 and 0.52, while maintaining equal weights among other dimensions. This can be viewed as an animation to examine change in relative index values as a response to the changing weights. This visualization technique, which presents a sequence of data projections, is referred to as a "tour" and the specific kind of tour used here to move between nearby projections is known as a "radial tour" (see @buja2005computational, @wickham_tourr_2011 and @RJ-2020-027 for more details).

### 2024-04-24 tidyindex.tex
Small changes and fixes

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This echoes an issue raised
in Donoho (2017), that different data analysts might arrive at different
conclusions despite using the same data. This is especially pertinent to
index use which affects important decisions such as in natural disaster
prevention, economic interventions, resource allocation or human
development. It is primarily due to a lack of standards in data analysis
workflow. Donoho (2017) called for research on structuring unified
workflow to address methodological variation across studies in data
science. Current practices also violate statistical principles, where
quantifying and understanding uncertainty are essential to deciding on a
best measure or metric, for example by incorporating bootstrap
confidence intervals (Efron 1979). There has been considerable research
in tidying up routine data analyses (Wickham 2014, 2011; Kuhn and Silge
2022; Wang, Cook, and Hyndman 2020; Zhang et al. 2024) that ``turn ideas
into software quickly and faithfully'', as envisioned by Chambers
(1998). Index development and use needs tidying.

@@

To be able to test different decision choices for an index,
systematically and statistically, the index needs to be broken down into
its fundamental building blocks to analyze the contribution and effect
of each component. We call this process the \emph{index pipeline}, which
are the steps of the data analysis pipeline for the index construction.
Such decomposition of index components provides the means to standardize
index construction via a pipeline and offers benefits for comparing
among indexes, calculating index uncertainty, and assessing index
robustness.

@@

Since the success of tidyverse, more packages have been developed to
analyze data using the tidy framework for domains specific applications,
a noticeable example of which is \texttt{tidymodels} for building
machine learning models (Kuhn and Silge 2022). To create a tidy workflow
tailored to a specific domain, developers first need to identify the
fundamental building blocks to create a workflow. These components are
then implemented as modules, which can be combined to form the pipeline.
For example, in supervised machine learning models, steps such as data
splitting, model training, and model evaluation are commonly used in
most workflow. In the \texttt{tidymodels}, these steps are
correspondingly implemented as package \texttt{rsample},
\texttt{parsnip}, and \texttt{yardstick}, agnostic to the specific model
chosen. The uniform interface in tidymodels frees analysts from
recalling model-specific syntax for performing the same operation across
different models, increasing the efficiency to work with different
models simultaneously.

@@

\begin{itemize}
\tightlist
\item
  makes the computation more transparent, and thus more easily debugged,
  facilitating reproducibility.
\item
  allows for rapidly processing new data to check how different
  features, like outliers, might affect the index value.
\item
  provides the capacity to measure uncertainty by computing confidence
  intervals from multiple samples as generated by bootstrapping to
  original data.
\item
  enables systematic comparison of surrogate indexes designed to measure
  the same phenomenon.
\item
  it may even be possible to automate diagrammatic explanations and
  documentation of the index.
\end{itemize}

@@

Unlike variable transformation, scaling maintains the distributional
shape of the variable. It includes techniques such as centering, z-score
standardization, and min-max standardization and can be expressed as
\([x_{j}(s;t) - \alpha]/\gamma\) where \(\alpha\) and \(\gamma\) are
constants . In the Human Development Index (HDI), the three dimensions
(health, education, and economy) are converted into the same scale (0-1)
using min-max standardization.

@@

The state of Queensland in Australia frequently experiences natural
disaster events such as flood and drought, which can significantly
impact its agricultural industry. This example uses daily data from
Global Historical Climatology Network Daily (GHCND), aggregated into
monthly precipitation, to compute two drought indexes -- SPI and SPEI --
at various time scales and fitted distributions, for 29 stations in the
state of Queensland in Australia, spanning from January 1990 to April
2022. This example showcases the basic calculation of indexes with
different parameter specifications within the pipeline framework. The
dataset used in this example is available in the \texttt{tidyindex}
package as \texttt{queensland} and blow prints the first few rows of the
data:

\begin{verbatim}
# A tibble: 5 x 9
  id                ym  prcp  tmax  tmin  tavg  long   lat name       
  <chr>          <mth> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <chr>      
1 ASN00029038 1990 Jan  1682  34.3  24.7  29.5  142. -15.5 KOWANYAMA ~
2 ASN00029038 1990 Feb   416  35.2  23.2  29.2  142. -15.5 KOWANYAMA ~
3 ASN00029038 1990 Mar  2026  32.5  23.6  28.0  142. -15.5 KOWANYAMA ~
4 ASN00029038 1990 Apr   597  32.9  17.7  25.3  142. -15.5 KOWANYAMA ~
5 ASN00029038 1990 May   244  31.8  20.1  25.9  142. -15.5 KOWANYAMA ~
\end{verbatim}

@@

\begin{verbatim}
Rows: 128,576
Columns: 18
$ .idx   <chr> "spei", "spei", "spei", "spei", "spei", "spei", "spei~
$ .dist  <chr> "gev", "gev", "gev", "gev", "gev", "gev", "gev", "gev~
$ id     <chr> "ASN00029038", "ASN00029038", "ASN00029038", "ASN0002~
$ month  <dbl> 6, 7, 8, 9, 10, 11, 12, 12, 1, 1, 2, 2, 3, 3, 4, 4, 5~
$ ym     <mth> 1990 Jun, 1990 Jul, 1990 Aug, 1990 Sep, 1990 Oct, 199~
$ prcp   <dbl> 170, 102, 0, 0, 0, 278, 1869, 1869, 5088, 5088, 8484,~
$ tmax   <dbl> 29.65357, 31.20323, 31.32581, 32.80870, 36.80357, 36.~
$ tmin   <dbl> 16.25000, 17.15161, 13.11613, 16.25714, 21.49655, 24.~
$ tavg   <dbl> 22.95179, 24.17742, 22.22097, 24.53292, 29.15006, 30.~
$ long   <dbl> 141.7483, 141.7483, 141.7483, 141.7483, 141.7483, 141~
$ lat    <dbl> -15.4818, -15.4818, -15.4818, -15.4818, -15.4818, -15~
$ name   <chr> "KOWANYAMA AIRPORT", "KOWANYAMA AIRPORT", "KOWANYAMA ~
$ .pet   <dbl> 67.46933, 86.64868, 63.27450, 94.93572, 204.63793, 24~
$ .diff  <dbl> 102.53067, 15.35132, -63.27450, -94.93572, -204.63793~
$ .scale <chr> "6", "6", "6", "6", "6", "6", "6", "12", "6", "12", "~
$ .agg   <dbl> 4263.7863, 2819.8773, 2529.0243, 578.8843, -117.6571,~
$ .fit   <dbl> 0.02902164, 0.10512807, 0.57680687, 0.83297600, 0.818~
$ .index <dbl> -1.89537090, -1.25286143, 0.19373133, 0.96599235, 0.9~
\end{verbatim}

@@

Figure~\ref{fig-idx-tour} illustrates doing sensitivity analysis for
GGGI, for a subset of 16 countries. It presents 6 frames selected from
the animation where the weight on the politics dimension is gradually
increased, where other dimensions (economy, education, health) decrease
correspondingly. Frame 12 present the original index where all the four
dimensions receive equal weight. The index values are sorted from
highest to lowest, with the Nordic countries (Iceland, Norway, and
Finland) and New Zealand leading the rankings. The index values are
between 0 and 1, and indicate proportional difference between men and
women, with a value of 0.8 indicating women are 80\% of the way to
equality of these measures. There is a gap in values between these
countries and the middle group (Brazil, Panama, Poland, Bangladesh,
Kazakhstan, Armenia, and Slovakia), and another big drop to the next
group (Pakistan, Iran, Algeria, and Chad). Afghanistan lags much further
behind.

@@

To make a simple illustration of sensitivity analysis, we slightly vary
the weight for politics, between 0.07 and 0.52, while maintain equal
weights among other dimensions. This can be viewed as an animation to
examine change in relative index values as a response to the changing
weights. This visualization technique, which presents a sequence of data
projections, is referred to as a ``tour'' and the specific kind of tour
used here to move between nearby projections is known as a ``radial
tour'' (see Andreas Buja et al. (2005), Wickham et al. (2011) and
Spyrison and Cook (2020) for more details).

>>>>

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This echoes an issue raised
in Donoho (2017), that different data analysts might arrive at different
conclusions despite using the same data. This is especially pertinent to
index use which affects important decisions such as in natural disaster
prevention, economic interventions, resource allocation or human
development. It is primarily due to a lack of standards in data analysis
workflow. Donoho (2017) called for research on structuring a unified
workflow to address methodological variation across studies in data
science. Current practices also violate statistical principles, where
quantifying and understanding uncertainty are essential to deciding on a
best measure or metric, for example by incorporating bootstrap
confidence intervals (Efron 1979). There has been considerable research
in tidying up routine data analyses (Wickham 2014, 2011; Kuhn and Silge
2022; Wang, Cook, and Hyndman 2020; Zhang et al. 2024) that ``turn ideas
into software quickly and faithfully'', as envisioned by Chambers
(1998). Index development and use needs tidying.

@@

To be able to test different decision choices for an index,
systematically and statistically, the index needs to be broken down into
its fundamental building blocks to analyze the contribution and effect
of each component. We call this process the \emph{index pipeline}, which
are the steps of the data analysis pipeline for index construction. Such
a decomposition of index components provides the means to standardize
index construction via a pipeline and offers benefits for comparing
among indexes, calculating index uncertainty, and assessing index
robustness. In addition it provides an unambiguous documentation for the
index definition, facilitating reproducibility of results.

@@

Since the success of tidyverse, more packages have been developed to
analyze data using the tidy framework for domains specific applications,
a noticeable example of which is \texttt{tidymodels} for building
machine learning models (Kuhn and Silge 2022). To create a tidy workflow
tailored to a specific domain, developers first need to identify the
fundamental building blocks to create a workflow. These components are
then implemented as modules, which can be combined to form the pipeline.
For example, in supervised machine learning models, steps such as data
splitting, model training, and model evaluation are commonly used in
most workflow. In the \texttt{tidymodels}, these steps are
correspondingly implemented as packages \texttt{rsample},
\texttt{parsnip}, and \texttt{yardstick}, agnostic to the specific model
chosen. The uniform interface in tidymodels frees analysts from
recalling model-specific syntax for performing the same operation across
different models, increasing the efficiency to work with different
models simultaneously.

@@

\begin{itemize}
\tightlist
\item
  makes the computation more transparent, and thus more easily debugged,
  facilitating reproducibility.
\item
  allows for rapidly processing new data to check how different
  features, like outliers, might affect the index value.
\item
  provides the capacity to measure uncertainty by computing confidence
  intervals from multiple samples as generated by bootstrapping the
  original data.
\item
  enables systematic comparison of surrogate indexes designed to measure
  the same phenomenon.
\item
  it may even be possible to automate diagrammatic explanations and
  documentation of the index.
\end{itemize}

@@

Unlike variable transformation, scaling maintains the distributional
shape of the variable. It includes techniques such as centering, z-score
standardization, and min-max standardization and can be expressed as
\([x_{j}(s;t) - \alpha]/\gamma\) where \(\alpha\) and \(\gamma\) are
constants. In the Human Development Index (HDI), the three dimensions
(health, education, and economy) are converted into the same scale (0-1)
using min-max standardization.

@@

The state of Queensland in Australia frequently experiences natural
disaster events such as flood and drought, which can significantly
impact its agricultural industry. This example uses daily data from
Global Historical Climatology Network Daily (GHCND), aggregated into
monthly precipitation, to compute two drought indexes -- SPI and SPEI --
at various time scales and fitted distributions, for 29 stations in the
state of Queensland in Australia, spanning from January 1990 to April
2022. This example showcases the basic calculation of indexes with
different parameter specifications within the pipeline framework. The
dataset used in this example is available in the \texttt{tidyindex}
package as \texttt{queensland} and below we show the first few rows of
the data:

\begin{verbatim}
# A tibble: 5 x 9
  id                ym  prcp  tmax  tmin  tavg  long   lat name       
  <chr>          <mth> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <chr>      
1 ASN00029038 1990 Jan  1682  34.3  24.7  29.5  142. -15.5 KOWANYAMA ~
2 ASN00029038 1990 Feb   416  35.2  23.2  29.2  142. -15.5 KOWANYAMA ~
3 ASN00029038 1990 Mär  2026  32.5  23.6  28.0  142. -15.5 KOWANYAMA ~
4 ASN00029038 1990 Apr   597  32.9  17.7  25.3  142. -15.5 KOWANYAMA ~
5 ASN00029038 1990 Mai   244  31.8  20.1  25.9  142. -15.5 KOWANYAMA ~
\end{verbatim}

@@

\begin{verbatim}
Rows: 128,576
Columns: 18
$ .idx   <chr> "spei", "spei", "spei", "spei", "spei", "spei", "spei~
$ .dist  <chr> "gev", "gev", "gev", "gev", "gev", "gev", "gev", "gev~
$ id     <chr> "ASN00029038", "ASN00029038", "ASN00029038", "ASN0002~
$ month  <dbl> 6, 7, 8, 9, 10, 11, 12, 12, 1, 1, 2, 2, 3, 3, 4, 4, 5~
$ ym     <mth> 1990 Jun, 1990 Jul, 1990 Aug, 1990 Sep, 1990 Okt, 199~
$ prcp   <dbl> 170, 102, 0, 0, 0, 278, 1869, 1869, 5088, 5088, 8484,~
$ tmax   <dbl> 29.65357, 31.20323, 31.32581, 32.80870, 36.80357, 36.~
$ tmin   <dbl> 16.25000, 17.15161, 13.11613, 16.25714, 21.49655, 24.~
$ tavg   <dbl> 22.95179, 24.17742, 22.22097, 24.53292, 29.15006, 30.~
$ long   <dbl> 141.7483, 141.7483, 141.7483, 141.7483, 141.7483, 141~
$ lat    <dbl> -15.4818, -15.4818, -15.4818, -15.4818, -15.4818, -15~
$ name   <chr> "KOWANYAMA AIRPORT", "KOWANYAMA AIRPORT", "KOWANYAMA ~
$ .pet   <dbl> 67.46933, 86.64868, 63.27450, 94.93572, 204.63793, 24~
$ .diff  <dbl> 102.53067, 15.35132, -63.27450, -94.93572, -204.63793~
$ .scale <chr> "6", "6", "6", "6", "6", "6", "6", "12", "6", "12", "~
$ .agg   <dbl> 4263.7863, 2819.8773, 2529.0243, 578.8843, -117.6571,~
$ .fit   <dbl> 0.02902164, 0.10512807, 0.57680687, 0.83297600, 0.818~
$ .index <dbl> -1.89537090, -1.25286143, 0.19373133, 0.96599235, 0.9~
\end{verbatim}

@@

Figure~\ref{fig-idx-tour} illustrates doing sensitivity analysis for
GGGI, for a subset of 16 countries. It presents 6 frames selected from
an animation where the weight on the politics dimension is gradually
increased, while other dimensions (economy, education, health) decrease
correspondingly. Frame 12 present the original index where all the four
dimensions receive equal weight. The index values are sorted from
highest to lowest, with the Nordic countries (Iceland, Norway, and
Finland) and New Zealand leading the rankings. The index values are
between 0 and 1, and indicate proportional difference between men and
women, with a value of 0.8 indicating women are 80\% of the way to
equality of these measures. There is a gap in values between these
countries and the middle group (Brazil, Panama, Poland, Bangladesh,
Kazakhstan, Armenia, and Slovakia), and another big drop to the next
group (Pakistan, Iran, Algeria, and Chad). Afghanistan lags much further
behind.

@@

To make a simple illustration of sensitivity analysis, we slightly vary
the weight for politics, between 0.07 and 0.52, while maintaining equal
weights among other dimensions. This can be viewed as an animation to
examine change in relative index values as a response to the changing
weights. This visualization technique, which presents a sequence of data
projections, is referred to as a ``tour'' and the specific kind of tour
used here to move between nearby projections is known as a ``radial
tour'' (see Andreas Buja et al. (2005), Wickham et al. (2011) and
Spyrison and Cook (2020) for more details).
